TacticToe: Learning the Game of Tactics

### 2018-01-21 tactictoe_jv.tex
search: non-constructive definition of a solved goal and node

\subsection{A proof tree with two branching factors}

A tactical proof tree is a set of nodes containing a list of goals where each 
goal $g$ of a node $n$ is linked to a node $m$ when the list of

@@

\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
\draw[-to,dotted,thick] (0) to (1m);
\draw[-to,dotted,thick] (0) to (2m);
\draw[-to,dotted,thick] (0) to (3b);
\end{tikzpicture}
\end{center}
\caption{The two branching factors: choice of a goal $g_0^i$ and choice of a 
tactic $t_j$.}
\label{fig:choice}
\end{figure}

Starting from the bottom node, we can construct descendant of this node by 
choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven.

@@

\paragraph{Solving a goal and a node}
Describe the case where a proof for a goal in a node is found.
If one of those list where empty, the goal $g_i^0$ would 
be proved.
A node is solved when every of its goal are solved.

@@

\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{Policy(a_i)}\]

\[Policy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The 
square root skews this probability to favor even more exploration of nodes with 
few visits.

The exploitation function $Exploitation$ is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree.   
\[Exploitation(a_i) = Average_{b \in Desc(a_i))} {PriorEvaluation(b)}\]

@@

tacictoe is resourceful.

@@

Original \holfour proof script (2.03 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
REWRITE_TAC [IMAGE_SURJ] THEN
  DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
  BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC
  THENL [
  irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
    REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
    REPEAT STRIP_TAC THEN REPEAT BasicProvers.VAR_EQ_TAC THEN
    FULL_SIMP_TAC std_ss [THE_DEF],
  ASM_REWRITE_TAC [LINV_OPT_def, o_THM, THE_DEF] THEN
    RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
      [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC [] ]
\end{lstlisting}

@@

\tactictoe proof script (50.85 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
\end{lstlisting}

@@

%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
%   `!l1 l2 n x.
%      LENGTH l1 <= n ==>
%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
%);
\vspace{5mm}

\tactictoe proof script (17.16 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
Induct_on `l1` THENL [SRW_TAC [] [],
  Cases_on `n` THENL [SRW_TAC [] [],
    FULL_SIMP_TAC (srw_ss ()) [] THEN METIS_TAC [LUPDATE_def]]]
\end{lstlisting}

If a discovered proof is reasonable short, fast, maintainable  and 
understandable, they could 
replace its original in the \holfour standard library. In the first case the 
tactictoe proof is slower but could still be considered becaues it is much more 
readable. In the second case, it is faster and more readable.

@@

Future work:
  Promising techniques that could not be evaluated. 
  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
  + Try harder and smarted on term predictions. \cite{latest version of sepia}. 
  \cite{conjecturing}

>>>>

\subsection{A search tree: data structures and definitions}

\begin{definition}(Tactical search tree)
A tactical search tree (or search tree) is a set of nodes containing a list of 
goals where each goal $g$ of a node $n$ is linked to a node $m$ when the list 
of 
\end{definition}

@@

\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
\draw[-to,dotted,thick] (0) to (1m);
\draw[-to,dotted,thick] (0) to (2m);
\draw[-to,dotted,thick] (0) to (3b);
\end{tikzpicture}
\end{center}
\caption{A node of a search tree and its relation with the children of one of 
its goal}
\label{fig:choice}
\end{figure}

Starting from the bottom parent node, we can construct children of a goal of 
this node by 
choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
apply to this goal. Here the result produces new nodes each containing a list 
of goals to be proven.

In each node can either be solved, open or pending. The definition of a solved 
goal is expressed through mutual recursion with a 
solved node and is given in ~\ref{def:solved_goal}. An open goal is goal that 
was tried at least once by a tactic. In our implementation we have only one 
open goals per node at once. A goal that was not explored yet is called a 
pending goal.

\begin{definition}(Solved node and solved goal)
Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^*$ and 
the set of solved goals $\mathbb{G}^*$ are the smallest set verifying the 
following equations:

\todo{should I make a constructive definition with $A^n+1$ in terms of $G^n$ 
instead?}
\begin{align*}
\mathbb{A}^* &= \lbrace a \in \mathbb{A}\ |\ \forall g \in G(a).\ g \in 
\mathbb{G}^* \rbrace\\
\mathbb{G}^* &= \lbrace g \in \mathbb{G}\ |\ \exists t \in T(g).\ t(g) \in 
\mathbb{A}^* \rbrace
\end{align*}

By the base case of these definition, a goal is solved if there exists a tactic 
applied to this goal that produces no goals.

@@

@@

\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]

\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
The policy is approximatively the  percentage of time a node was visited.The 
square root skews this probability to favor even more exploration of nodes with 
few visits.

The exploitation term is the average of the evaluation of 
all nodes descendant of $a_i$ (itself included) in the current search tree. It 
is also called the current evaluation of $a_i$.   
\[Exploitation(a_i) = CurEvaluation(a_i) = 
  Average_{b \in Desc(a_i)} {PriorEvaluation(b)}\]

@@

TacicToe is resourceful. Changing the shape of the feature of the theorems
to get better predictions.

@@

Original \holfour proof script (2.03 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
REWRITE_TAC [IMAGE_SURJ] THEN
DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC THENL 
  [irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
     REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
     REPEAT STRIP_TAC THEN REPEAT BasicProvers.VAR_EQ_TAC THEN
     FULL_SIMP_TAC std_ss [THE_DEF],
   ASM_REWRITE_TAC [LINV_OPT_def, o_THM, THE_DEF] THEN
     RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
       [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC []]
\end{lstlisting}

@@

Make it match the new version.
\tactictoe proof script (50.85 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
\end{lstlisting}

@@

%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
%   `!l1 l2 n x.
%      LENGTH l1 <= n ==>
%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
%);
\vspace{5mm}
Make it match the new version.
\tactictoe proof script (17.16 milliseconds)
\begin{lstlisting}[language=SMLSmall,frame=tb]
Induct_on `l1` THENL [SRW_TAC [] [],
  Cases_on `n` THENL [SRW_TAC [] [],
    FULL_SIMP_TAC (srw_ss ()) [] THEN METIS_TAC [LUPDATE_def]]]
\end{lstlisting}
fs () is short for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.

If a discovered proof is short, fast, maintainable and 
understandable, it could replace its original in the \holfour standard library.

@@

Future work:
  Promising techniques that could not be evaluated. 
  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
  + Try harder and smarted on term predictions. \cite{latest version of sepia}. 
  \cite{conjecturing}
  + Self-learning.  
+ MCTS was presented in manner that we only did the first step of reinforcement 
learning. CurrentPolicy becomes PriorPolicy.
